---
title: "1_acquire.qmd"
format: html
---
```{r}
#| label: loading packages
#| message: false
library(dplyr) # data manipulation
library(readr) # data import/ export
library(gutenbergr) # Project Gutenberg API
library(skimr)      # descriptive statistics
library(knitr)      # tables
library(ggplot2)    # plotting
```

## Description of data

The data I will be using comes from Project Gutenberg, which, according to [Recipe 5](https://qtalr.github.io/qtalrkit/articles/recipe-5.html) is a "volunteer effort to digitize and archive cultural works." Most fo the works in the database are public domain in the US which means they can be freely accessed, used, and shared. I downloaded the package gutenbergr which will allow me to access the metadata for all of the works available in their database. 

```{r}
glimpse(gutenberg_metadata)
```
This data shows that there are 72,569 works in Project Gutenberg's database. 

#### I am going to save this data as a csv file 

```{r}
metadata_df <-gutenberg_metadata

write_csv(metadata_df, "data/original/gutenberg_metadata.csv")
```

When I loaded this lab, the instructions said to open a quarto document under the process directory. This is that document. However, when I went to write my data and save it as a CSV file, I ran into some issues. I wanted to save the file under `data/original` but it keps saving under the process directory and I couldnt figure out why. I think this is because that both `process` and `data` are under the `lab-05` directory together. I didnt know how to make the file save under `lab-05/data/original` so I created a new data folder under the process directory. I was then able to save my CSV file under `lab-05/process/data/original` but since I was already working on a document in the process directory, I just had to write `data/original` in the code. I deleted the data folder under the `lab-05` directory.

#### I am also interested in the authors, so I will save metadata about the Gutenberg authors

```{r}
glimpse(gutenberg_authors)

authors_df <-gutenberg_authors

write_csv(authors_df, "data/original/gutenberg_authors.csv")
```

#### I will do the same for the subjects of each work

```{r}
glimpse(gutenberg_subjects)

subjects_df <-gutenberg_subjects

write_csv(subjects_df, "data/original/gutenberg_subjects.csv")
```

## Deriving data 

```{r}

metadata_read <- read_csv("data/original/gutenberg_metadata.csv")
write_csv(metadata_read, "data/derived/gutenberg_metadata_read.csv")

authors_read <- read_csv("data/original/gutenberg_authors.csv")
write_csv(authors_read, "data/derived/gutenberg_authors_read.csv")

subjects_read <- read_csv("data/original/gutenberg_subjects.csv")
write_csv(subjects_read, "data/derived/gutenberg_subjects_read.csv")

```

Ok that might have been useless because I'm looking at `authors_df` and `authors_read` and I think they are just the same thing. I think maybe what I was trying to do was summarize the data. I'll try that instead now to display the results

## Summary of results

```{r}
summary(authors_read)
```

```{r}
summary(metadata_read)
```

```{r}
summary(subjects_read)
```
